---
title: "IS-2 Donation Model"
author: "Clarice Muthoni Gitonga"
date: "17/11/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |                 |
|----------------------------------------------|-----------------|
| **Student ID Number**                        | 127707          |
| **Student Name**                             | Clarice Gitonga |
| **BBIT 4.2 Group**                           | 4.2B            |
| **Supervisor**                               | Selina Evelia   |

# Introduction to Milestone 3

## 1: Install and use renv
```{r}
renv::init()
```

## 2: Loading the dataset
```{r}
library(readr)
census <- read_csv("data/census.csv")
View(census)
```

Reduce datasets
```{r}
census2 <- census[sample(nrow(census), 1000), ]
View(census2)
```

Save as CSV
```{r}
write.csv(census2, file = "C:/Users/clari/OneDrive/Desktop/API/BIProject/data/census2.csv", row.names = FALSE)
```

drop the native-country dataset
```{r}
census2 <- subset(census2, select = -`native-country`)
```


## 3. Install and Load the Required Packages
```{r}
## caret
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## klaR
if (require("klaR")) {
  require("klaR")
} else {
  install.packages("klaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## LiblineaR
if (require("LiblineaR")) {
  require("LiblineaR")
} else {
  install.packages("LiblineaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## naivebayes
if (require("naivebayes")) {
  require("naivebayes")
} else {
  install.packages("naivebayes", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## rpart
if (require("rpart")) {
  require("rpart")
} else {
  install.packages("rpart", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}


## MASS
if (require("MASS")) {
  require("MASS")
} else {
  install.packages("MASS", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## glmnet
if (require("glmnet")) {
  require("glmnet")
} else {
  install.packages("glmnet", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## stats
if (require("stats")) {
  require("stats")
} else {
  install.packages("stats", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## corrplot
if (!is.element("corrplot", installed.packages()[, 1])) {
  install.packages("corrplot", dependencies = TRUE)
}
require("corrplot")

## ggcorrplot
if (!is.element("ggcorrplot", installed.packages()[, 1])) {
  install.packages("ggcorrplot", dependencies = TRUE)
}
require("ggcorrplot")

## Amelia
if (!is.element("Amelia", installed.packages()[, 1])) {
  install.packages("Amelia", dependencies = TRUE)
}
require("Amelia")

## dplyr - For data manipulation
if (!is.element("dplyr", installed.packages()[, 1])) {
  install.packages("dplyr", dependencies = TRUE)
}
require("dplyr")

## ggplot2 - For data visualizations using the Grammar for Graphics package
if (!is.element("ggplot2", installed.packages()[, 1])) {
  install.packages("ggplot2", dependencies = TRUE)
}
require("ggplot2")

## ggrepel - Additional options for the Grammar for Graphics package
if (!is.element("ggrepel", installed.packages()[, 1])) {
  install.packages("ggrepel", dependencies = TRUE)
}
require("ggrepel")

## ggraph - Additional options for the Grammar for Graphics package
if (!is.element("ggraph", installed.packages()[, 1])) {
  install.packages("ggraph", dependencies = TRUE)
}
require("ggraph")

## tidytext - For text mining
if (!is.element("tidytext", installed.packages()[, 1])) {
  install.packages("tidytext", dependencies = TRUE)
}
require("tidytext")

## tidyr - To tidy messy data
if (!is.element("tidyr", installed.packages()[, 1])) {
  install.packages("tidyr", dependencies = TRUE)
}
require("tidyr")

## widyr - To widen, process, and re-tidy a dataset
if (!is.element("widyr", installed.packages()[, 1])) {
  install.packages("widyr", dependencies = TRUE)
}
require("widyr")

## gridExtra - to arrange multiple grid-based plots on a page
if (!is.element("gridExtra", installed.packages()[, 1])) {
  install.packages("gridExtra", dependencies = TRUE)
}
require("gridExtra")

## knitr - for dynamic report generation
if (!is.element("knitr", installed.packages()[, 1])) {
  install.packages("knitr", dependencies = TRUE)
}
require("knitr")

## kableExtra - for nicely formatted output tables
if (!is.element("kableExtra", installed.packages()[, 1])) {
  install.packages("kableExtra", dependencies = TRUE)
}
require("kableExtra")

## formattable -  To create a formattable object
# A formattable object is an object to which a formatting function and related
# attributes are attached.
if (!is.element("formattable", installed.packages()[, 1])) {
  install.packages("formattable", dependencies = TRUE)
}
require("formattable")

## circlize - To create a cord diagram or visualization
# by Gu et al. (2014)
if (!is.element("circlize", installed.packages()[, 1])) {
  install.packages("circlize", dependencies = TRUE)
}
require("circlize")

## memery - For creating data analysis related memes
# The memery package generates internet memes that optionally include a
# superimposed inset plot and other atypical features, combining the visual
# impact of an attention-grabbing meme with graphic results of data analysis.
if (!is.element("memery", installed.packages()[, 1])) {
  install.packages("memery", dependencies = TRUE)
}
require("memery")

## magick - For image processing in R
if (!is.element("magick", installed.packages()[, 1])) {
  install.packages("magick", dependencies = TRUE)
}
require("magick")

## yarrr - To create a pirate plot 
if (!is.element("yarrr", installed.packages()[, 1])) {
  install.packages("yarrr", dependencies = TRUE)
}
require("yarrr")

## radarchart - To create interactive radar charts using ChartJS 
if (!is.element("radarchart", installed.packages()[, 1])) {
  install.packages("radarchart", dependencies = TRUE)
}
require("radarchart")

## igraph - To create ngram network diagrams 
if (!is.element("igraph", installed.packages()[, 1])) {
  install.packages("igraph", dependencies = TRUE)
}
require("igraph")

## wordcloud2 - For creating wordcloud by using 'wordcloud2.JS 
if (!is.element("wordcloud2", installed.packages()[, 1])) {
  install.packages("wordcloud2", dependencies = TRUE)
}
require("wordcloud2")

## textdata - Download sentiment lexicons and labeled text data sets 
if (!is.element("textdata", installed.packages()[, 1])) {
  install.packages("textdata", dependencies = TRUE)
}
require("textdata")

## stringr - For processing characters in a string
if (!is.element("stringr", installed.packages()[, 1])) {
  install.packages("stringr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
require("stringr")

## pROC
if (require("pROC")) {
  require("pROC")
} else {
  install.packages("pROC", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```

# Algorithm Selection for Classification
## A. Linear Algorithms
### 1. Generalized Linear Model
Define a 70:30 train:test data split of the dataset.
```{r}
train_index <- createDataPartition(census2$income,
                                   p = 0.7,
                                   list = FALSE)
census_train <- census2[train_index, ]
census_test <- census2[-train_index, ]
```

#### Train the Model
We apply the 5-fold cross validation resampling method
```{r}
train_control <- trainControl(method = "cv", number = 5)

set.seed(7)
income_model_glm <-
  train(income ~ ., data = census_train, method = "glm",
        metric = "Accuracy", trControl = train_control)
```

#### Display the Model's Performance 
```{r}
print(income_model_glm)
summary(income_model_glm)
str(census_test[, 1:13])
str(census_train[, 1:13])
```

#### Make predictions
```{r}
predictions <- predict(income_model_glm, census_test[, 1:12])

new_levels <- setdiff(levels(census_test$`marital-status`), levels(income_model_glm$terms$`marital-status`))

# Check the new levels
print(new_levels)

# Set the levels of workclass in the testing dataset to match the training dataset
census_test$`marital-status` <- factor(census_test$`marital-status`, levels = levels(income_model_glm$terms$`marital-status`))

# Print the first few rows of the predictions
print(predictions)

```

#### Display the model's evaluation metrics
```{r}
confusion_matrix <-
  caret::confusionMatrix(predictions,as.factor(
    census_test[, 1:13]$income))

print(confusion_matrix)

fourfoldplot(as.table(confusion_matrix), color = c("grey", "lightblue"),
             main = "Confusion Matrix")
```

