---
title: "IS-2 Donation Model"
author: "Clarice Muthoni Gitonga"
date: "17/11/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |                 |
|----------------------------------------------|-----------------|
| **Student ID Number**                        | 127707          |
| **Student Name**                             | Clarice Gitonga |
| **BBIT 4.2 Group**                           | 4.2B            |
| **Supervisor**                               | Selina Evelia   |

# Introduction to Milestone 1

## 1: Install and use renv
```{r}
renv::init()
```

## 2: Loading the dataset
```{r}
library(readr)
census <- read_csv("data/census.csv")
View(census)
```

## 3: Preview Loaded Dataset
### Dimensions 
Because the dataset is large and has 45222 rows and 14 columns, it will be reduced to 1000 rows and 14 columns.
```{r}
dim(census)
```

Reduce datasets
```{r}
census2 <- census[sample(nrow(census), 1000), ]
View(census2)
```

Save as CSV
```{r}
write.csv(census2, file = "C:/Users/clari/OneDrive/Desktop/API/BIProject/data/census2.csv", row.names = FALSE)
```

### Data Types
```{r}
sapply(census2, class)
```

# Descriptive Statistics
## A. Measures of frequency
```{r}
census2_freq <- census2$income   # For Income
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$relationship   # For Relationship
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$occupation      # For Occupation
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$workclass        # For workclass
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$education_level   # For education_level
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$`marital-status`   # For marital_status
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$sex   # For sex
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$race   # For race
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

census2_freq <- census2$`native-country`   # For native_country
cbind(frequency = table(census2_freq),
      percentage = prop.table(table(census2_freq)) * 100)

```

## B.Measures of Central Tendency
Calculate the Mode
```{r}
# For income
census2_income_mode <- names(table(census2$income))[      
  which(table(census2$income) == max(table(census2$income)))
]
print(census2_income_mode)

# For age
census2_age_mode <- names(table(census2$age))[      
  which(table(census2$age) == max(table(census2$age)))
]
print(census2_age_mode)
```

## C. Measure of distribution of the data
### 1. Summary of Distribution
```{r}
summary(census2)
```

### 2. Standard deviation
```{r}
sapply(census2[,c(1,4,10,11,12)], sd)
```

### 3. Measure of Variance
```{r}
sapply(census[,c(1,4,10,11,12)], var)
```

### 4. Measure of Kurtosis
capital-gain, capital-loss and hours-per-week have a high number of outliers.
```{r}
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")

sapply(census[,c(1,4,10,11,12)], kurtosis, type = 2)
```

### 5. Measure of Skewness 
Skewness above 0.4 implies a positive skew; a right-skewed distribution. (age, capital-gain, capital-loss)

Skewness below -0.4 implies a negative skew; a left-skewed distribution.(hours-per-week)
```{r}
sapply(census[,c(1,4,10,11,12)], skewness, type = 2)
```

## D. Measures of Relationship
### 1. Measure the covariance between variables
```{r}
census2_cov <- cov(census2[, c(1,4,10,11,12)])
View(census2_cov)
```

### 2. Measure the correlation between variables
```{r}
census2_cor <- cor(census2[, c(1,4,10,11,12)])
View(census2_cor) 
```

# Inferential Statistics
### ANOVA
Dependent variable:	Income.
Independent variables: age, workclass, education_level, education-num, marital-status, occupation, relationship, race, sex, capital-gain, capital-loss, hours-per-week, native-country.

#### (i) One-Way ANOVA
```{r}
result_one_way <- aov(`hours-per-week` ~ education_level, data = census2)
summary(result_one_way)
```

#### (ii) Two-Way ANOVA
```{r}
result_two_way <- aov(`hours-per-week` ~ education_level * `marital-status`, data = census2)
summary(result_two_way)
```

# Basic Visualizations
## (i) Univariate Plots
### Histograms for Each Numeric Attribute
```{r}
# Age
age <- as.numeric(unlist(census2[, 1]))
hist(age, main = names(census2)[1])

# Education-num
`education-num` <- as.numeric(unlist(census2[, 4]))
hist(`education-num`, main = names(census2)[4])

# Capital-gain
`capital-gain` <- as.numeric(unlist(census2[, 10]))
hist(`capital-gain`, main = names(census2)[10])

# Capital-loss
`capital-loss` <- as.numeric(unlist(census2[, 11]))
hist(`capital-loss`, main = names(census2)[11])

# hours-per-week
`hours-per-week` <- as.numeric(unlist(census2[, 12]))
hist(`hours-per-week`, main = names(census2)[12])
```

### Box and Whisker Plots for Each Numeric Attribute
```{r}
boxplot(census2[, 1], main = names(census2)[1])
boxplot(census2[, 4], main = names(census2)[4])
boxplot(census2[, 10], main = names(census2)[10])
boxplot(census2[, 11], main = names(census2)[11])
boxplot(census2[, 12], main = names(census2)[12])
```

### Bar Plots for Each Categorical Attribute
```{r}
barplot(table(census2[, 2]), main = names(census2)[2])
barplot(table(census2[, 3]), main = names(census2)[3])
barplot(table(census2[, 5]), main = names(census2)[5])
barplot(table(census2[, 6]), main = names(census2)[6])
barplot(table(census2[, 7]), main = names(census2)[7])
barplot(table(census2[, 8]), main = names(census2)[8])
barplot(table(census2[, 9]), main = names(census2)[9])
barplot(table(census2[, 13]), main = names(census2)[13])
barplot(table(census2[, 14]), main = names(census2)[14])
```

## (ii) Multivariate Plots
```{r}
# ggplot
if (!is.element("ggcorrplot", installed.packages()[, 1])) {
  install.packages("ggcorrplot", dependencies = TRUE)
}
require("ggcorrplot")

# caret
if (!is.element("caret", installed.packages()[, 1])) {
  install.packages("caret", dependencies = TRUE)
}
require("caret")

# listenv
if (!requireNamespace("listenv", quietly = TRUE)) {
  install.packages("listenv", dependencies = TRUE)
}
library(listenv)

# parallelly
if (!requireNamespace("parallelly", quietly = TRUE)) {
  install.packages("parallelly", dependencies = TRUE)
}
library(parallelly)

# globals
if (!is.element("globals", installed.packages()[, 1])) {
  install.packages("globals", dependencies = TRUE)
}
library(globals)
```

#### Correlation Plot
```{r}
ggcorrplot(cor(census2[,c(1,4,10,11,12)]))
```

#### Create a Scatter Plot ----
```{r}
pairs(~ age  + `education-num` + `hours-per-week`, data = census2, col = as.numeric(factor(census2$income)))

# Alternatively, the ggcorrplot package can be used to make the plots more
# appealing:
ggplot(census2,
       aes(x = `hours-per-week`, y = `marital-status`, shape = income, color = income)) +
  geom_point() +
  geom_smooth(method = lm)
```

### Multivariate Box and Whisker Plots by Class
```{r}
featurePlot(x = census2[, c(1,4,10,11,12)], y = census2[, 14], plot = "box")
```


