---
title: "IS-2 Donation Model"
author: "Clarice Muthoni Gitonga"
date: "18/11/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |                 |
|----------------------------------------------|-----------------|
| **Student ID Number**                        | 127707          |
| **Student Name**                             | Clarice Gitonga |
| **BBIT 4.2 Group**                           | 4.2B            |
| **Supervisor**                               | Selina Evelia   |

# Introduction to Milestone 2

## 1: Install and use renv
```{r}
renv::init()
```

## 2: Loading the dataset
```{r}
library(readr)
census <- read_csv("data/census.csv")
View(census)

# Reduce datasets
census2 <- census[sample(nrow(census), 1000), ]
View(census2)

# Save as CSV
write.csv(census2, file = "C:/Users/clari/OneDrive/Desktop/API/BIProject/data/census2.csv", row.names = FALSE)

```

# Missingness Map to Identify Missing Data
There was no missing data.
```{r}
# Amelia 
if (!is.element("Amelia", installed.packages()[, 1])) {
  install.packages("Amelia", dependencies = TRUE)
}
require("Amelia")

missmap(census2, col = c("red", "grey"), legend = TRUE)
```

# Data Transforms
Data transforms can improve the accuracy of your final model when applied as part of the pre-processing stage.
```{r}
## Install and Load the Required Packages
## mlbench
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## factoextra
if (require("factoextra")) {
  require("factoextra")
} else {
  install.packages("factoextra", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## FactoMineR
if (require("FactoMineR")) {
  require("FactoMineR")
} else {
  install.packages("FactoMineR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```

## Basic data transforms
### (i) Scale Data Transform
```{r}
# BEFORE
summary(census2)

# Age
age <- as.numeric(unlist(census2[, 1]))
hist(age, main = names(census2)[1])

# Education-num
`education-num` <- as.numeric(unlist(census2[, 4]))
hist(`education-num`, main = names(census2)[4])

# Capital-gain
`capital-gain` <- as.numeric(unlist(census2[, 10]))
hist(`capital-gain`, main = names(census2)[10])

# Capital-loss
`capital-loss` <- as.numeric(unlist(census2[, 11]))
hist(`capital-loss`, main = names(census2)[11])

# hours-per-week
`hours-per-week` <- as.numeric(unlist(census2[, 12]))
hist(`hours-per-week`, main = names(census2)[12])

model_of_the_transform <- preProcess(census2, method = c("scale"))
print(model_of_the_transform)
census2_scale_transform <- predict(model_of_the_transform,
                                          census2)

# AFTER
summary(census2_scale_transform )

# Age
hist(census2_scale_transform$age, main = names(census2_scale_transform)[which(names(census2_scale_transform) == 'age')])

# Education-num
hist(census2_scale_transform$`education-num`, main = names(census2_scale_transform)[which(names(census2_scale_transform) == 'education-num')])

# Capital-gain
hist(census2_scale_transform$`capital-gain`, main = names(census2_scale_transform)[which(names(census2_scale_transform) == 'capital-gain')])

# Capital-loss
hist(census2_scale_transform$`capital-loss`, main = names(census2_scale_transform)[which(names(census2_scale_transform) == 'capital-loss')])

# hours-per-week
hist(census2_scale_transform$`hours-per-week`, main = names(census2_scale_transform)[which(names(census2_scale_transform) == 'hours-per-week')])
```

### (ii) Centering Data Transform
```{r}
summary(census2)
model_of_the_transform <- preProcess(census2, method = c("center"))
print(model_of_the_transform)
census2_center_transform <- predict(model_of_the_transform, census2)
summary(census2_center_transform)
```

### (iii) Standardization
```{r}
summary(census2)

# BEFORE
sapply(census[,c(1,4,10,11,12)], sd)
model_of_the_transform <- preProcess(census2,
                                     method = c("scale", "center"))
print(model_of_the_transform)
census2_standardize_transform <- predict(model_of_the_transform, census2)
# AFTER
summary(census2_standardize_transform)
sapply(census2_standardize_transform[,c(1,4,10,11,12)], sd)

```

### (iv) Normalization
```{r}
summary(census2)
model_of_the_transform <- preProcess(census2, method = c("range"))
print(model_of_the_transform)
census2_normalize_transform <- predict(model_of_the_transform, census2)
summary(census2_normalize_transform)
```

## Power transforms
### (i)Yeo-Johnson
```{r}
# BEFORE
summary(census2)

#Calculate the skewness before the Box-Cox transform
sapply(census2[,c(1,4,10,11,12)], skewness, type = 2)

#histogram to view the skewness before the Box-Cox transform
# Age
age <- as.numeric(unlist(census2[, 1]))
hist(age, main = names(census2)[1])

# Education-num
`education-num` <- as.numeric(unlist(census2[, 4]))
hist(`education-num`, main = names(census2)[4])

# Capital-gain
`capital-gain` <- as.numeric(unlist(census2[, 10]))
hist(`capital-gain`, main = names(census2)[10])

# Capital-loss
`capital-loss` <- as.numeric(unlist(census2[, 11]))
hist(`capital-loss`, main = names(census2)[11])

# hours-per-week
`hours-per-week` <- as.numeric(unlist(census2[, 12]))
hist(`hours-per-week`, main = names(census2)[12])

model_of_the_transform <- preProcess(census2, method = c("YeoJohnson"))
print(model_of_the_transform)
census2_yeo_johnson_transform <- predict(model_of_the_transform, # nolint
                                                census2)
# AFTER
summary(census2_yeo_johnson_transform)

# Calculate the skewness after the Yeo-Johnson transform
sapply(census2_yeo_johnson_transform[,c(1,4,10,11,12)],  skewness, type = 2)

# Age
hist(census2_yeo_johnson_transform$age, main = names(census2_yeo_johnson_transform)[which(names(census2_yeo_johnson_transform) == 'age')])

# Education-num
hist(census2_yeo_johnson_transform$`education-num`, main = names(census2_yeo_johnson_transform)[which(names(census2_yeo_johnson_transform) == 'education-num')])

# Capital-gain
hist(census2_yeo_johnson_transform$`capital-gain`, main = names(census2_yeo_johnson_transform)[which(names(census2_yeo_johnson_transform) == 'capital-gain')])

# Capital-loss
hist(census2_yeo_johnson_transform$`capital-loss`, main = names(census2_yeo_johnson_transform)[which(names(census2_yeo_johnson_transform) == 'capital-loss')])

# hours-per-week
hist(census2_yeo_johnson_transform$`hours-per-week`, main = names(census2_yeo_johnson_transform)[which(names(census2_yeo_johnson_transform) == 'hours-per-week')])
```

## Linear algebra transforms
### (i) Principal Component Analysis (PCA)
```{r}
summary(census2)

# Exclude 'income' from the variables being scaled
model_of_the_transform <- preProcess(census2[, -14], method = c("scale", "center", "pca"))

print(model_of_the_transform)
census2_pca_dr <- predict(model_of_the_transform, census2)

summary(census2_pca_dr)
dim(census2_pca_dr)
```

#### PCA for Feature Extraction
```{r}
census2_pca_fe <- princomp(cor(census2[,c(1,4,10,11,12)]))
summary(census2_pca_fe)
```

#### Scree Plot
```{r}
factoextra::fviz_eig(census2_pca_fe, addlabels = TRUE)

```

```{r}
census2_pca_fe$loadings[, 1:2]
```


```{r}
# it shows the extent to which each variable is represented 
factoextra::fviz_cos2(census2_pca_fe, choice = "var", axes = 1:2)
```

#### Biplot and Cos2 Combined Plot
```{r}
factoextra::fviz_pca_var(census2_pca_fe, col.var = "cos2",
                         gradient.cols = c("brown", "orange", "green"),
                         repel = TRUE)
```

### (ii) Independent Component Analysis (ICA)
```{r}
# fastICA
if (!is.element("fastICA", installed.packages()[, 1])) {
  install.packages("fastICA", dependencies = TRUE)
}
require("fastICA")

summary(census2)

model_of_the_transform <- preProcess(census2,
                                     method = c("scale", "center", "ica"),
                                     n.comp = 1)
print(model_of_the_transform)
census2_ica_dr <- predict(model_of_the_transform, census2)

summary(census2_ica_dr)

```

